---
title: "FinalR"
author: "Miao Chen"
date: "December 16, 2015"
output: html_document
---

###CMDS on Delta matrix from word-topic matrix

First, we read in the data from .csv file

```{r}
#read in the .csv file into R. It was read as a data frame.
DocTopicData=read.csv("C:\\Users\\Miao\\Desktop\\Courses\\S675 Statistical Learning\\topics20.csv",header=TRUE)
#show the first sevearl rows of the document-topic matrix
head(DocTopicData)
```

Compute the pairwise dissimilarities:
```{r}
#use the 2nd to last columns in the dataframe, because the 1st column is the text file name; and then convert it to a Matrix, X.
X <- data.matrix(DocTopicData[,-1])
#get squre root for the matrix elements. The rows represent points on teh unit sphere.
Z <- sqrt(X) 
#generate the (cosine) similarity matrix
Gamma <- Z %*% t(Z)
#generate the dissimilarity betweenrows i and j of X
Delta <- acos(Gamma)
#make the diagonal elements as 0, as they were NaN due to arccos computation
diag(Delta) <- 0
```

```{r}
Delta.cmds <- cmdscale(Delta, k=4, eig=TRUE)
X <- Delta.cmds$points
```

Let r denote the number of strictly positive eigenvalues. Then r=348 according to the following command.
```{r}
length(subset(Delta.cmds$eig, Delta.cmds$eig>0))
```


```{r}
#sum of the positive eigenvalues
sum.positive_eig=sum(subset(Delta.cmds$eig, Delta.cmds$eig>0))
sum.positive_eig

#sum of the negative eigenvalues
sum.negative_eig=sum(subset(Delta.cmds$eig, Delta.cmds$eig<0))
sum.negative_eig

#percentage of variation capture by the first k=4 principal components:
sum(Delta.cmds$eig[1:4])/sum.positive_eig
```

The first k=4 dimensional representation captures 53.62% of the total variation.

```{r}
sum(Delta.cmds$eig[1:4])/sum.positive_eig
```

The amount of variation captured by d=4 dimensions out of the total variation is 53.62%. This seems not so sufficent because there is still 45.38% variation not captured, which is very close to the amount being captured. 

```{r}
#get all the positive eigenvalues
positive_eigvalues=subset(Delta.cmds$eig, Delta.cmds$eig>0)
#get cumulative sum for this vector
cumsum_eigenvalues=cumsum(positive_eigvalues)
#then get % of variation captured by each d<=r dimension
variation.percentage=cumsum(positive_eigvalues)/sum.positive_eig
#then plot the variation percentage at different d<=r dimension
plot(variation.percentage, xlab="num of dimensions")
#here we also show the captured variations of the first 24 dimension's 
variation.percentage[1:24]
```



###CMDS on KL divergence

```{r}
Doc_Doc_KLDistance=read.csv("C:\\Users\\Miao\\Desktop\\Courses\\S675 Statistical Learning\\final\\darwin-voyage\\KLs\\kl_dists20.csv",header=TRUE)
#show the first sevearl rows of the document-topic matrix
head(Doc_Doc_KLDistance)[,1:10]

#use the 2nd to last columns in the dataframe, because the 1st column is the text file name; and then convert it to a Matrix, X.
Delta1 <- data.matrix(Doc_Doc_KLDistance[,-1])

Delta1.cmds <- cmdscale(Delta1, k=4, eig=TRUE)

#sum of the positive eigenvalues
sum.positive_eig=sum(subset(Delta1.cmds$eig, Delta1.cmds$eig>0))
sum.positive_eig

#sum of the negative eigenvalues
sum.negative_eig=sum(subset(Delta1.cmds$eig, Delta1.cmds$eig<0))
sum.negative_eig

#percentage of variation capture by the first k=4 principal components:
sum(Delta1.cmds$eig[1:4])/sum.positive_eig
```

```{r}
length(subset(Delta1.cmds$eig, Delta1.cmds$eig>0))
```

```{r}
#get all the positive eigenvalues
positive_eigvalues=subset(Delta1.cmds$eig, Delta1.cmds$eig>0)
#get cumulative sum for this vector
cumsum_eigenvalues=cumsum(positive_eigvalues)
#then get % of variation captured by each d<=r dimension
variation.percentage=cumsum(positive_eigvalues)/sum.positive_eig
#then plot the variation percentage at different d<=r dimension
plot(variation.percentage, xlab="num of dimensions")
#here we also show the captured variations of the first 24 dimension's 
variation.percentage[1:250]

